Conversion of FFCS ICE fleet to EV: a data-driven approach

### 2020-11-17 Chapter5/5_1_introduction.tex
added_chap_10
abstract added

\section{Introduction}
The aim of this work is to find a methodology to convert combustion engine FFCS into electric one using real data. In order to do that,  I developed an event-driven simulator able to replicate the users' FFCS mobility patterns created from the data collected with the tools described in chapters \ref{chap:2_dataset} and \ref{chap:4_cs_comparison}.

In order to give a main idea, the simulator takes as input a real trace composed by a ordered set of rentals, an operative area composed by adjacent squares zones of 0.025 $km_2$, the set of charging station and their placement, a car model and a fleet size. By consuming the trace, the simulator computes trip by trip the amount of energy needed to travel the proper distance and moves the designed car from the starting point to the chosen destination.

During the simulations, the software computes several metrics in order to measure to proper size the charging infrastructure and how it is refleceted as user discomfort, i.e. in terms of number of plugging operation.

>>>>

\section{Introduction}
In the previous chapters I widely described the dataset in I collected and characterized. However the collected data refers \emph{only} to internal combustion engine cars. Recalling that the main research question of my thesis is: \emph{It is possible design an electric Free Floating Car Sharing System?}, I need a tool able to replicate the FFCS customers' habits in a customizable model of electric FFCS system.

In this chapter I describe the core-tool of my research: an event-driven trace-based electric FFCS simulator. In a nutshell, this software I developed is able to (i)  extract users pattern from data collected in chapter \ref{chap:2_dataset}; (ii) replicate the users patterns in a electrified scenario parametrized by the client; (iii) return several performance metrics describing the inputted electric FFCS setup.

In order to give a main idea, the simulator takes as input a real trace composed by a ordered set of rentals. More in details, the simulator creates the event-trace composed by \emph{Event} events. Each one of them carries temporal and spatial coordinates (when the event triggers) and it can represent a rental start or rental end. In this way, the simulator is able to mirrors the exact FFCS trips.

Another input is the operative area composed by adjacent squares zones of 0.025 $km^2$. Then, each one of those zone can be equipped with a charging station, so the set of charging station and their placement is the third input. Complete the input set a car model and a fleet size.

By consuming the trace, the simulator computes trip by trip the amount of energy needed to travel the proper distance and moves the designed car from the starting point to the chosen destination.

During the simulations, the software computes several metrics in order to measure to proper size the charging infrastructure and how it is reflected as user discomfort, i.e. in terms of number of plugging operation.

### 2020-11-17 Chapter5/5_3_placement.tex
abstract added

\section{Meta-Heuristic Charging Stations Placement}
\label{sec:5_3_mh_placement}
In this sections I explain the ratio behind the charging station placement. Indeed, in the previous section I described the structure and simulation policies, while now I define an important environmental variable that will be deeply studied in chapters \mc{references}. I remind that from the rental booking trace, I extrapolate the service area and I divide it into square zones of 500 m of side length. Then, using the number of parking or the average parking time for each zone, the simulator choose which zones equip with a charging station.

@@

This brief catheterization shows how  different cities can have different spatial characterization and thus different charging station placements. However those characterization will be deepened in chapters \mc{captoli}.

>>>>

\section{Meta-Heuristic Charging Stations Placement}
\label{sec:5_3_mh_placement}
In this sections I explain the charging station placement algorithm. The output of this algorithm is one of the most relevant environmental variables that will be deeply studied and analysed in chapters \ref{chap:6_4cities} and \ref{chap:7_cs_optimization}. The main idea behind this algorithm to rank each zone (defined in section \ref{sec:5_2_modelling}) according a the users' travelling patterns and, then equip zones having the highest values.

@@

This brief catheterization shows how  different cities can have different spatial characterization and thus different charging station placements. However those characterization will be deepened in chapter \ref{chap:6_4cities}.

### 2020-11-17 Chapter5/5_4_charging_policies.tex
abstract added

\section{Users' Plugging Policy}
\label{sec:5_4_return_policy}
One of the most challenging points of electric FFCS is to deal with the discomfort derived by plug in operations. This operation is more time consuming with compared to the normal filling up procedure of combustion engine cars. Therefore, the providers have to deal with users' selfishness and trying to stimulate their willingness.  
In this Section, we define different policies that the electric FFCS may enforce, and different probabilistic behaviors of customers.

\subsection{Car charging policies}
When returning the car, the customer may connect the car to a pole in a station, hence charging the car battery and possibly deviating the real destination from the desired one. I modelled the following policies:
\begin{itemize}
	\item{\it Free Floating}: the customer must connect the car to a charging pole if and only if it is available in the desired final zone $D(i)$;
	\item{\it Forced}: cars must be connected to a pole when the percentage of battery charge at the end of the rental $i$ would go below a certain threshold $\pi$, i.e., $(c(a,t_{s}(i)) - Energy(p(a,t_{s}(i)), d(i))) \cdot 100/ C\leq  \pi $. This implies the customer can be \textit{rerouted} to the closest zone to the desired one $d(i)$, if no free pole exists in the zone; %Battery consumption takes into account the additional traveled distance.
	\item{\it Hybrid}: the customers follow the forced policy; they may  also choose to connect to a charging pole available in the desired ending zone $D(i)$ with probability $w\in [0,1]$;
	%if the $c(a,t)\leq\pi$, cars must be returned to the closest recharging zone to $d(i)$.
\end{itemize}

>>>>

\section{Car return policies}
\label{sec:5_4_return_policy}
One of the most challenging points of electric FFCS is to deal with the discomfort derived by plug in operations. This operation is more time consuming with compared to the normal filling up procedure of combustion engine cars. Therefore, the providers have to deal with users' selfishness and trying to stimulate their willingness.  
%In this Section, we define different policies that the electric FFCS may enforce, and different probabilistic behaviors of customers.

%\subsection{Car charging policies}
When returning the car, the customer may connect the car to a pole in a station, hence charging the car battery and possibly deviating the real destination from the desired one. I modelled the following policies:
\begin{itemize}
	\item{\it Free Floating}: the customer must connect the car to a charging pole if and only if it is available in the desired final zone $D(i)$;
	\item{\it Forced}: cars must be connected to a pole when the percentage of battery charge at the end of the rental $i$ would go below a certain threshold $\pi$, i.e., $(c(a,t_{s}(i)) - Energy(p(a,t_{s}(i)), d(i))) \cdot 100/ C\leq  \pi $. This implies the customer can be \textit{rerouted} to the closest zone to the desired one $d(i)$, if no free pole exists in the zone; %Battery consumption takes into account the additional traveled distance.
	\item{\it Hybrid}: the customers follow the forced policy; they may  also choose to connect to a charging pole available in the desired ending zone $D(i)$ with probability $w\in [0,1]$;
	%if the $c(a,t)\leq\pi$, cars must be returned to the closest recharging zone to $d(i)$.
\end{itemize}

### 2020-11-17 thesis.tex
abstract added

\summary%\sommario
%This is where you write your abstract \dots\ (Maximum 4000 characters, i.e. maximum two pages in normal sized font, typeset with the thesis layout).
%The abstract environment is also available, but  \texttt{\string\summary} is preferred because it generates an un-numbered chapter. The abstract environment is more suitable for articles and two column typesetting without a separate title page.
\mc{intro}

@@

Then, I characterized both datasets in Turin, one of the cities in which both FFCS providers work. Initially, I defined a filtering criterion to remove outliers from the dataset and identify only real users' ride: entries reporting a trip where the user realistically moved. These analyses allowed me to identify geo-spatial user's patterns. After that, I compared the car2go customer's pattern with the one-way and two-way car-sharing system. The results showed how users prefer more flexible services like FFCS or one-way car sharing.

@@

At this point I studied in Berlin, Milan, Turin, and Vancouver (cities having different traffic patterns) different electric FFCS setup. The best logic to place a charging station is to cover the zones recording more parkings: increasing the charging station presence in those zones, the probability to find a free plug increases too. Then, I showed how an electric FFCS is sustainable with few charging stations (8-10~\% of city zones) if the users accept to park the car in a charging station when their ride finishes nearby or when the SoC goes below a security threshold. Finally, I investigate different charging station setups. The best results come by balancing the spread of charging stations: concentrated all plugs in a unique hub or spread them all above the city leads to poor performances in terms of system sustainability and user's discomforts.

To still reduce the number of charging stations to have a sustainable electric FFCS, I compared several optimization algorithms. My results showed how a Genetic Algorithm is able to find a better solution able to shrink the minimum amount of resources to sustain the same mobility demand.

After that, I moved my attention to the users' rentals predictability. The main goal is to understand how different open-data sources could impact on the recorded FFCS users' rental. I used as a case of study the city of Vancouver, augmenting the rentals dataset with the weather and socio-economic open data. Initially, I compared several time-series forecasts to predict the users' demand in the short and medium-term. The Random Forrest regression produced better accuracy and results interpretability. Then I correlated the socio-economics features characterizing each Vancouver neighborhood to FFCS demand and again, the Random Forrest regression clearly showed a subset of socio-economic features more correlated to the FFCS users' rentals.

>>>>

\summary%\sommario
%This is where you write your abstract \dots\ (Maximum 4000 characters, i.e. maximum two pages in normal sized font, typeset with the thesis layout).
%The abstract environment is also available, but  \texttt{\string\summary} is preferred because it generates an un-numbered chapter. The abstract environment is more suitable for articles and two column typesetting without a separate title page.
Nowadays, the increase of traffic congestions, land consume and pollution emission due to a private car ownership opened the way to share mobility. One of the most spread implementation of shared mobility is Free Floating Car Sharing (FFCS). It is a car rental model where the users can pick and release the car everywhere within an operative area. The customers can reserve (and release) the vehicle through web-based application. With just a simple tap, the users can unlock and lock the smart vehicles. Usually the provider bills the users only for the time spend driving, with time-minute based fares. All the other costs like petrol, insurance and maintenance are in charge to the provider.

The flexibility of this service fills the urban mobility offer gap between public transport and private car ownership. Indeed, FFCS allows people to travel and commute faster than common public bus but avoiding all the fixed and variable costs related to private car ownership.

Given the recent increase market of the electric vehicles and all the pros those vehicles carry, replace FFCS fleet with electric powered cars may still improve the quality of life in urban centres. The setup and management of an electric FFCS requires ingenuity, in order to minimize the users discomfort due the plugging procedures .

@@

Then, I characterized both datasets in Turin, one of the cities in which both FFCS providers work. I defined outlier detection filtering criteria and I extracted geo-temporal users' travel patterns.

After that, I compared the car2go customer's pattern with the one-way and two-way car-sharing system. The results showed how users prefer more flexible services like FFCS or one-way car sharing.

@@

%At this point I studied in Berlin, Milan, Turin, and Vancouver (cities having different traffic patterns) different electric FFCS setup. I showed that placing the charging station where the users ends the rental most frequently and makes the system sustainable with few charging stations (8-10~\% of city zones). Finally, I investigate different charging station setups. The best results come by balancing the spread of charging stations.
%concentrated all plugs in a unique hub or spread them all above the city leads to poor performances in terms of system sustainability and user's discomforts.

Then, via accurate simulation in Berlin, Milan, Turin and Vancouver, I studied different electric FFCS setup. By placing the charging station in the most frequented areas, incentive the users to plug the car when the battery state of charge is below a safety threshold and balancing the spread of poles, it is possible to obtain a sustainable system covering with chaging station only the 8-10~\% of zones

To still reduce the number of charging stations to have a sustainable electric FFCS, I compared several optimization algorithms. The results showed how a Genetic Algorithm is able to find a better solution able to shrink the minimum amount of resources to sustain the same mobility demand.

After that, I moved my attention to the users' rentals predictability. The main goal is to understand how different open-data sources could impact on the recorded FFCS users' rental. Initially, I compared several time-series forecasts to predict the users' demand in the short and medium-term. The Random Forrest regression produced better accuracy and results interpretability. Then I correlated the socio-economics features characterizing each Vancouver neighborhood to FFCS demand and again, the Random Forrest regression outperforms other algorithms
